Legends for supplementary material

### 2019-02-01 Manuscript/Figures/figure_legends.Rmd
editted legends

**Figure 3. The three predictor types considered when quantifying the association between TAD boundaries and genomic annotations.** The _overlap count_ (OC) predictors consider the total number of genomic features that overlapped with each genomic bin. The _overlap percent_ (OP) predictors were calculated by dividing the sum of all feature widths within each specific bin by the total bin width. The _distance_ predictors were calculated by measuring the distance (in bases) from the center of each genomic bin to the center of the nearest genomic feature.

**Figure 4. Precision-Recall curves for unbalanced 10 kb data.** Precision-recall curves were used to initially compare performance of random forest models with different predictor types on unbalanced data at 10 kb resolution. GM12878 (solid line) and K562 (dashed line) cell lines were compared with predictor type-specific models which included overlap counts (blue), overlap percents (purple), log distance (green), and all three combined (red). For both cell lines, random forest models with log distance-type predictors were found to have higher AUPRC values.

**Figure 5. Model Performances on 10kb data using distance-type predictors for different number of features** Using RFE, performances of a series of random forest models with different numbers of the most predictive features for each cell line was compared using AUPRC. For convenience and computational efficiency, only specific subsets of features were considered (increasing powers of 2). For the 10 kb data, near optimal performance was achieved for both the GM12878 (red) and K562 cell lines (blue).

>>>>

**Figure 3. The three predictor types considered when quantifying the association between TAD boundaries and genomic annotations.** The _overlap count_ (OC) predictors consider the total number of genomic features that overlapped with each genomic bin. The _overlap percent_ (OP) predictors were calculated by dividing the sum of all feature widths within each specific bin by the total bin width. The _distance_ ($log_{2}$ transformed) predictors were calculated by measuring the distance (in bases) from the center of each genomic bin to the center of the nearest genomic feature.

**Figures 4-8. Precision-Recall curves for 10 kb data.** Precision-recall curves were used to initially compare performance of random forest models with different predictor types on 10 kb resolution data, across each resampling technique. GM12878 (solid line) and K562 (dashed line) cell lines were compared with predictor type-specific models which included overlap counts (blue), overlap percents (purple), log distance (green), and all three combined (red).

**Figure 9-13. Recursive Feature Elimination for 10 kb data** Beginning with the full set of distance type predictors for each cell line, models were recursively evaluated at different subsets of feature size. For convenience and computational efficiency, feature size decreasing in powers of 2 was considered. For the 10 kb data, near optimal performance was achieved for both the GM12878 (red) and K562 cell lines (blue). The maximum size for feature subsets differed between cell line due the available annotations provided in Supplementary Table 1.

Performances of a series of random forest models with different numbers of the most predictive features for each cell line was compared using AUPRC. For convenience and computational efficiency, only specific subsets of features were considered (increasing powers of 2).

Additionally, models using the full feature space were also evaluated, hence the differing number of features between cell lines. For the 10 kb data, near optimal performance was achieved for both the GM12878 (red) and K562 cell lines (blue).

### 2019-02-01 Manuscript/Supplementary/supplementary_legends.Rmd
editted legends

**Figure 1. Distance Densities across cell lines and resolutions**
Density curves of distances (in millions of base pairs) from genomic bins to flanked TAD boundaries for GM12878 (left column) and K562 (right column) cell lines. The untransformed distances (red) are highly right skewed. After performing a $log_{2}$ tranformation (blue), the distances exhibit more normalized densities. The normalized distances were exhibited across all data resolution (rows).

**Figure 2. Precision-Recall curves for unbalanced 25 kb, 50kb, and 100kb data.** Precision-recall curves were used to initially compare performance of random forest models with different predictor types on unbalanced data at 25 kb, 50kb, and 100kb resolution. GM12878 (solid line) and K562 (dashed line) cell lines were compared with predictor type-specific models which included overlap counts (blue), overlap percents (purple), log distance (green), and all three combined (red). For both cell lines, and among all resolutions, random forest models with log distance-type predictors were found to have higher AUPRC values.

@@

**Supplementary Table S1. List of Genomic Annotations.** The GRCh37/hg19 human genome assembly was used. "Genomic Class" - broad category of genomic features, "Element" - names of genomic features, "Cell Type" - cell type, "Source" - download URL. [??? Did you use only features common to both cell lines?]

>>>>

**Supplementary Table 4. Hi-C Data.** A complete list of the Hi-C data libraries used to obtain the locations of topologically associating domains and where they can be downloaded.

**Supplementary Table 1. Genomic Annotations.** A complete list of the genomic annotations used to build the feature space for all downstream models. The GRCh37/hg19 human genome assembly was used. "Genomic Class" - broad category of genomic features, "Element" - names of genomic features, "Available Cell Type" - the name of the cell type in which the annotation was available, "Cell Type-Specific Source" - download URL specific to the cell line (not all annotations were provided by the same institutions). [??? Did you use only features common to both cell lines?; !!!clarified, need to discuss further]

**Supplementary Figure 1-8. Distance Densities across cell lines and resolutions.**
Density curves of distances (in millions of base pairs) from genomic bins to flanked TAD boundaries for GM12878 (left column) and K562 (right column) cell lines. The untransformed distances (red) are highly right skewed. After performing a $log_{2}$ tranformation (blue), the distances exhibited more normalized densities.

**Supplementary Table 5. Data Summaries and Imbalance Ratios.** The total number of genomic bins included in the complete, training, and testing sets, as well as the subsequent imbalance ratios across resolutions and cell lines. Imbalance ratio is defined as the ratio of genomic bins that overlapped a flanked TAD boundary (minority class), divided by the number of genomic bins that did not overlap a flanked TAD boundary (majority class).

**Supplementary Figures 9-23. Precision-Recall curves for unbalanced 25 kb, 50kb, and 100kb data.** Precision-recall curves were used to initially compare performance of random forest models with different predictor types on 25 kb, 50kb, and 100kb resolution data, across each resampling technique. GM12878 (solid line) and K562 (dashed line) cell lines were compared with predictor type-specific models which included overlap counts (blue), overlap percents (purple), log distance (green), and all three combined (red). The baseline threshold for each PR curve is defined as the ratio of genomic bins that overlapped with a flanked TAD boundary divided by the total number of genomic bins for the test set, respective to cell line and resolution.

**Supplementary Table 2. Area under Precision-Recall Curves.** Comparison of models applied to different data resolutions, resampling techniques, and predictor types for each of the two cell lines in terms of AUPRC. The column labeled "All" denotes models with all 3 combined predictor types.

@@
